---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Bri Rick"
date: "2-10-2022"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```


# Simple web scraping

R can read html using either rvest, xml, or xml2 packages. Here we are going to navigate to the Center for Snow and Avalance Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data we want to programatically download for three sites. We don't know much about these sites, but they contain incredibly rich snow, temperature, and precip data. 


## Reading an html 

### Extract CSV links from webpage

```{r, warning=FALSE, message=FALSE}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#See if we can extract tables and get the data that way
tables <- webpage %>%
  html_nodes('table') %>%
  magrittr::extract2(3) %>%
  html_table(fill = TRUE)
#That didn't work, so let's try a different approach

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('24hr',.)] %>%
  html_attr('href')

```

## Data Download

### Download data in a for loop

```{r, warning=FALSE, message=FALSE}

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('data/',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


### Download data in a map

```{r, warning=FALSE, message=FALSE}

#Map version of the same for loop (downloading 3 files)
if(evaluate == T){
  map2(links[1:3],file_names[1:3],download.file)
}else{print('data already downloaded')}

```

## Data read-in 

### Read in just the snow data as a loop

```{r, warning=FALSE, message=FALSE}
#Pattern matching to only keep certain files
snow_files <- file_names %>%
  .[!grepl('SG_24',.)] %>%
  .[!grepl('PTSP',.)]

#empty_data <- list()

# snow_data <- for(i in 1:length(snow_files)){
#   empty_data[[i]] <- read_csv(snow_files[i]) %>%
#     select(Year,DOY,Sno_Height_M)
# }

#snow_data_full <- do.call('rbind',empty_data)

#summary(snow_data_full)
```


### Read in the data as a map function

```{r, message=FALSE, warning=FALSE}

our_snow_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_24hr.csv','',.)
  df <- read_csv(file) %>%
    select(Year,DOY,Sno_Height_M) %>%
    mutate(site = name)
}

snow_data_full <- map_dfr(snow_files,our_snow_reader)

summary(snow_data_full)
```


### Plot snow data

```{r, warning=FALSE, message=FALSE}
snow_yearly <- snow_data_full %>%
  group_by(Year,site) %>%
  summarize(mean_height = mean(Sno_Height_M,na.rm=T))

ggplot(snow_yearly,aes(x=Year,y=mean_height,color=site)) + 
  geom_point() +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```


# Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.

```{r, warning=FALSE, message=FALSE}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')
```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r, warning=FALSE, message=FALSE}
#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('data/',dataset)

#Map version of the same for loop (downloading 2 files)
if(evaluate == T){
  map2(links[1:2],file_names[1:2],download.file)
}else{print('data already downloaded')}

```


3. Write a custom function to read in the data and append a site column to the data. 

```{r, warning=FALSE, message=FALSE}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:14] %>%
  str_trim(side = "left")

# Combine the files with forcing in them 
forcing_files <- file_names %>%
  .[grepl('Forcing',.)] 

# Create a function to read in data table, assign column names, and add site column
our_forcing_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_Forcing_Data.txt','',.)
  df <- read_table(file, col_names = headers, skip = 4) %>% select(c(1:14)) %>% 
    mutate(site = name)
}

```

4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.

```{r, warning=FALSE, message=FALSE}
# Use the map function to read in both of the forcing files. 
forcing_data_full <- map_dfr(forcing_files,our_forcing_reader) 

# Display a summary of the tibble
summary(forcing_data_full)
```


5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

*The average air temp for 2003 and 2004 look abnormally low and are excluded from the plot.*

```{r, warning=FALSE, message=FALSE}
# Create dataset with mean annual air temp per site
mean_yr_site <- forcing_data_full %>% filter(year > 2004) %>% group_by(year, site) %>% 
  summarise(meanK = mean(`air temp [K]`))

# Create a line plot of mean annual temp per site
ggplot(mean_yr_site, aes(x = year, y = meanK, color = site))+
  geom_line()+
  scale_x_continuous(breaks = c(2005, 2006, 2007, 2008, 2009, 2010, 2011))+
  ylab("Mean air temp [K]")+
  xlab("Year")+
  theme_bw()+
    theme(legend.position = c(.5, .85))
```


6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html

```{r, warning=FALSE, message=FALSE}
#Create a function that plots the monthly mean air temp per site
plotter <- function(yr){
    yrdata <- forcing_data_full %>% filter(year == yr) %>% 
    group_by(month, site) %>% 
    summarise(meanT = mean(`air temp [K]`))  
    print(ggplot(yrdata, aes(x = month, y = meanT, color = site)) + 
          geom_line() +
          theme_bw() + 
          theme(legend.position = c(.1, .8))+
          ylab("Mean air temp [K]")+
          xlab("Month")+
          ggtitle(yr))
}

#Use a for loop to plot monthly mean air temp per site for 2005 to 2010
for (i in 2005:2010){
        plotter(i)}

# Plots using just a for loop without creating a function
# for (i in 2005:2010){
#   yrdata <- forcing_data_full %>% filter(year == i) %>%
#     group_by(month, site) %>%
#     summarise(meanT = mean(`air temp [K]`))
#   print(ggplot(yrdata, aes(x = month, y = meanT, color = site)) +
#           geom_line() +
#           theme_bw() +
#           theme(legend.position = c(.1, .8))+
#           ylab("Mean air temp [K]")+
#           xlab("Month")+
#           ggtitle(i))
# }
```

Bonus: Make a plot of average daily precipitation by day of year (averaged across all available years). Color each site. 

Bonus #2: Use a function and for loop to create yearly plots of precipitation by day of year. Color each site. 
